Implementing simplified INLA into `aghq`

### 2022-10-21 src/dev_sinla/sinla.Rmd
Working through Wood algorithm

## Complete algorithm

In this section we omit dependence on $\theta$ such that: $\mu = \mu(\theta)$, $Q = Q(\theta)$, $\mu(x_i) = \mu(x_i, \theta)$ and $Q(x_i) = Q(x_i, \theta)$.

Want: $\tilde p(x_i \, | \, \theta, y)$

Let $\mathcal{D}_{2j}$ denote a $2j \times 2j$ diagonal matrix with leading diagonal $-1, 1, -1, 1, \ldots$.
Let $u_0$ be a zero column matrix.
Let $\Delta_j^i$ denote $\Delta_j$ with an extra zero inserted at element $i$.

1. Compute the Cholesky factor of $Q_0 := Q_{-i, -i}$ by update of the Cholesky factor of $Q$
2. For each $x_i$ in a grid of evaluation points (how does one chose this grid?) repeat the following steps a-e)
  a) Use Newton's method with fixed Hessian $Q_{-i, -i}$ to find $\mu(x_i)$ starting from \@ref(eq:conditionalmean)
  b) Compute a set of $J$ steps $\{\Delta_j\}$
  c) For $j = 1, \ldots, J - 1$ compute
  $$
  q = Q_0 \Delta_j + u_j \mathcal{D}_{2j}u_j^\top \Delta_j
  $$
  and
  $$
  g = \Delta_x \log p(\mu(x_i) + \Delta_j^i, y, \theta).
  $$
  Then compute the matrix
  $$
  u_{j + 1} = \{q(\Delta_j^\top j)^{-1/2}, g_{-i}(\Delta_j^\top g_{-i})^{-1/2}, u_j\}.
  $$
  d) Compute the determinant approximation...
  e) Compute $\tilde p(x_i \, | \, \theta, y) = \ldots$
3. Renormalise $\tilde p(x_i \, | \, \theta, y)$

>>>>

We follow Wood (2019) Section 2.
In this section we omit dependence on $\theta$ such that: $\mu = \mu(\theta)$, $Q = Q(\theta)$, $\mu(x_i) = \mu(x_i, \theta)$ and $Q(x_i) = Q(x_i, \theta)$.

Let $\mathcal{D}_{2j}$ denote a $2j \times 2j$ diagonal matrix with leading diagonal $-1, 1, -1, 1, \ldots$.
For example, with $j = 3$ we have:

```{r}
D2j <- function(j) {
  diag((-1)^{1:2 * j}, nrow = 2 * j)
}

D2j(3)
```

Let $u_0$ be a zero column matrix.
I am guessing that this means $u_j$ is a $j$ column matrix.
It's not specified what dimension it is too, but since it's multiplied by $\mathcal{D}_{2j}$ presumably it must be of length $2j$.

```{r}
uj <- function(j) {
 matrix(j, nrow = 2 * j, ncol = 1)
}

uj(3)
```

```{r}
D2j(3) %*% uj(3)
```

Let $\Delta_j^i$ denote $\Delta_j$ with an extra zero inserted at element $i$.
Note that it hasn't been defined what $\Delta_j$ is yet.

## Step 1: Compute the Cholesky factor

The first task is to "compute the Cholesky factor of $Q_0 := Q_{-i, -i}$ by update of the Cholesky factor of $Q$".

```{r}
i <- 1

mm <- h$env$last.par
mean <- mm[h$env$random]
Q <- h$env$spHess(mm, random = TRUE)
Q0 <- Q[-i, -i]

L0 <- chol(Q0)
```

We have that $Q_0 = L_0 {L_0}^\top$:

```{r}
sum(abs((L0 %*% t(L0)) - Q0)) < 10E-9
```

## Step 2. Loop

For each $x_i$ in a grid of evaluation points (how does one chose this grid?) repeat the following:

```{r}
spacing <- 0.05
x_grid <- seq(from = 0.5, to = 3.5, by = spacing)
```

### a)

Use Newton's method with fixed Hessian $Q_{-i, -i}$ to find $\mu(x_i)$ starting from \@ref(eq:conditionalmean):

```{r}
#' Conditional mean for the first grid point
Sigma <- solve(Q)
conditional_mean <- mean[-i] + Sigma[-i, i] * Sigma[i, i]^{-1} * (x_grid[1] - mean[i])
```

Note: for this model the latent field is IID, so the conditional mean is the same as the mean.

### b)

Compute a set of $J$ steps $\{\Delta_j\}$

```{r}
J <- 1
```

### c)

For $j = 1, \ldots, J - 1$ compute
$$
q = Q_0 \Delta_j + u_j \mathcal{D}_{2j}u_j^\top \Delta_j
$$
and
$$
g = \Delta_x \log p(\mu(x_i) + \Delta_j^i, y, \theta).
$$
Then compute the matrix
$$
u_{j + 1} = \{q(\Delta_j^\top j)^{-1/2}, g_{-i}(\Delta_j^\top g_{-i})^{-1/2}, u_j\}.
$$

### d)

Compute the determinant approximation
$$
\det(Q_1) = \det(Q_0) \det(I_{2J} + u_J^\top Q_0^{-1} u_J \mathcal{D}_{2j}).
$$

### e)

Compute $\tilde p(x_i \, | \, \theta, y) = \ldots$.

## Step 3. Renormalise $\tilde p(x_i \, | \, \theta, y)$

Renormalise $\tilde p(x_i \, | \, \theta, y)$.
